MSTK: Mesh Toolkit, v 1.1 - DRAFT

### 2003-11-17 docs/MSTK.tex
Added PDF and PS files to the repository Removed the .aux and .log files

\title{MSTK: Mesh Toolkit, v 1.0}

@@

\thispagestyle{empty}

\section{MSTK IO File Format}

\subsection{ASCII File Format}

>>>>

\title{MSTK: Mesh Toolkit, v 1.1 - DRAFT}

@@

\thispagestyle{empty}
\setlength{\parindent}{0.0in}
\setlength{\parskip}{0.5em}

\newpage
\section{Introduction}

MSTK or Mesh Toolkit is a mesh framework that allows users to
represent, manipulate and query unstructured 3D arbitrary topology
meshes in a general manner without the need to code their own data
structures. MSTK is a flexible framework in that it allows (or will
eventually allow) a wide variety of underlying representations for the
mesh while maintaining a common interface. It will allow users to
choose from different mesh representations either at initialization or
during the program execution so that the optimal data structures are
used for the particular algorithm. The interaction of users and
applications with MSTK is through a functional interface that acts as
though the mesh always contains vertices, edges, faces and regions and
maintains connectivity between all these entities.

\par MSTK allows for the simultaneous existence of an arbitrary number of
meshes. However, any entity in MSTK can belong to only one mesh at a
time.

\par MSTK will eventually support distributed meshes for parallel
computing. However, this is still not in place.

\par MSTK will soon allow applications to attach application or field data
to entities. This data may be integers, reals (doubles), integer
vectors, real (double) vectors, integer tensors, real (double tensors)
and pointers.

\par The basis for development of MSTK is laid out in the following paper:

Garimella, R. ``Mesh Data Structure Selection for Mesh Generation and
FEA Applications,'' \textit{International Journal of Numerical Methods
  in Engineering}, v55 n4, pp. 441-478, 2002.

In the following sections, the data types of MSTK will be described
followed by a description of the functional interface. The MSTK file
format will be described in the last section.

\newpage
\section{MSTK Data Types}

\begin{description}
\item[]\textbf{\textit{Set\_ptr}}: Handle to a Set object

\item[]\textbf{\textit{Mesh\_ptr}}: Handle to a Mesh object.

\item[]\textbf{\textit{MVertex\_ptr}}: Handle to a Mesh Vertex object (Topological
Dimension 0)

\item[]\textbf{\textit{MEdge\_ptr}}: Handle to a Mesh Edge object (Topological Dimension 1)

\item[]\textbf{\textit{MFace\_ptr}}: Handle to a Mesh Face object (Topological Dimension 2)

\item[]\textbf{\textit{MRegion\_ptr}}: Handle to Mesh Region object (Topological Dimension 3)

\item[]\textbf{\textit{MEntity\_ptr}}: Handle to a generic Mesh Entity object. Any of
the above types of entities can be cast as \textit{MEntity\_ptr}

\item[]\textbf{\textit{GModel\_ptr}}: Handle to a Geometric Model object
\item[]\textbf{\textit{GEntity\_ptr}}: Handle to a Geometric Entity object

\item[]\textbf{\textit{RepType}}: Enumerated type describing the type
  of mesh representation.  Can be UNKNOWN\_REP, F1, F4, R1, R2, R4.
  See Appendix~\ref{app:reptypes} for schematics of these
  representations \textit{Currently only representation types F1 and
    F4 are supported}

\item[]\textbf{\textit{MFType}}: Enumerated type for mesh face type
Can be FDELETED, FUNKNOWN, TRI, QUAD, POLYGON

\item[]\textbf{\textit{MRType}}: Enumerated type for mesh region type
Can be RDELETED, RUNKNOWN, TET, PYRAMID, PRISM, HEX, POLYHED
\end{description}

\newpage
\section{MSTK Functional Interface}
\subsection{Sets}

Sets of entities in MSTK are returned as sets of type
\textit{Set\_ptr}. The following are the set operations available in
MSTK:

\begin{description}
\item[]\textbf{\textit{Set\_ptr} Set\_New(\textit{int} inisize):} Create a
new set with an initial size, \textit{inisize}. If \textit{inisize} is
0, the initial size is set to be 10.

\item[]\textbf{\textit{void} Set\_Delete(\textit{Set\_ptr} l):} Delete a set.

\item[]\textbf{\textit{Set\_ptr} Set\_Compress(\textit{Set\_ptr} l):} Compress
a set.  Doing this while an algorithm is iterating through the set can
currently cause problems!! Calling \textbf{Set\_Compress} could change the
pointer for the set due to reallocation.

\item[]\textbf{\textit{Set\_ptr} Set\_Copy(\textit{Set\_ptr} l):} Return a
copy of a set.

\item[]\textbf{\textit{Set\_ptr} Set\_Add(\textit{Set\_ptr} l, \textit{void
  *}entry):} Add an entry to the set. The entry is appended to the end
of the set.

\item[]\textbf{\textit{Set\_ptr} Set\_ChknAdd(\textit{Set\_ptr} l,
\textit{void *}entry):} Add an entry to a set only if it is not already
in the set.

\item[]\textbf{\textit{int} Set\_Rem(\textit{Set\_ptr} l, \textit{void
  *}entry):} Remove an entry from the set. Returns 1 if successful, 0
otherwise.

\item[]\textbf{\textit{int} Set\_Remi(\textit{Set\_ptr} l, \textit{int} i):}
Remove the i'th valid entry in the set. Returns 1 if successful, 0
otherwise.

\item[]\textbf{\textit{int} Set\_Replace(\textit{Set\_ptr} l, \textit{void
  *}entry, \textit{void *}nuentry):} Replace 'entry' with 'nuentry' in
set. Returns 1 if successful, 0 otherwise.

\item[]\textbf{\textit{int} Set\_Replacei(\textit{Set\_ptr} l, \textit{int}
i, void *nuentry):} Replace the i'th valid entry in the set with
'nuentry'. Returns 1 if successful, 0 otherwise.

\item[]\textbf{\textit{int} Set\_Contains(\textit{Set\_ptr} l, void *entry):}
Returns 1 if set contains the entry, 0 otherwise.

\item[]\textbf{\textit{int} Set\_Locate(\textit{Set\_ptr} l, void *entry):}
Returns the positional index of the entry in the set. Returns -1 if
the set does not contain the entry.

\item[]\textbf{\textit{void *}Set\_Entry(\textit{Set\_ptr} l, \textit{int}
i):} Return the i'th valid entry in the set.  Returns a NULL pointer if
the i'th valid entry could not be found.

\item[]\textbf{\textit{void *}Set\_Next\_Entry(\textit{Set\_ptr} l,
\textit{int} *i):} Return the next valid entry in the set. This routine
works like an iterator. To start iterating through the set, set the
iteration index i=0 and call the routine to get the first entry in the
set. Subsequent calls to the routine will iterate through the entries
in the set. The routine will return a NULL to indicate that the end of
the set is reached.

The value of the iteration index i will be modified by the routine on
each call to indicate where in the set it is. This value should not be
modified externally while iterating through the set. Also, no specific
meaning be derived from from the iteration index by other applications
since the internal implementation and interpretation of the index may
change at any time.

\item[]\textbf{\textit{int} Set\_Num\_Entries(\textit{Set\_ptr} l):} Return
the number of entries in a set
\end{description}

\newpage
\subsection{Mesh Object}

A mesh object is a set of vertices (nodes) possibly connected by other
entities such as edges, faces, regions. Depending on the
representation chosen and type of mesh, some or all of the entities
may be explicitly stored.  Full representations contain all types of
entities up to the highest dimension of the mesh. For example, a full
representation of a tetrahedral mesh contains vertices, edges, faces
and regions. However, one type of reduced representation of this mesh
may contain only vertices and regions. For a surface mesh, a full
representation includes vertices, edges and faces while a reduced
representation only has vertices and faces. Also, depending on the
type of representation, some adjacencies (information about which
entities are connected to which other entities) are stored and others
are derived.

\begin{description}
\item[]\textbf{\textit{Mesh\_ptr} MESH\_New(\textit{RepType} type):}
Initialize a new mesh object with the given representation type which
can be F1, F2, F3, F4, F5, F6, R1, R2, R3, R4. Not all of these types
are implemented. If the representation type is not known at the
present time (e.g. before reading the mesh from a file), the
representation type of UNKNOWN\_REP can be specified.  Note that this
only initializes a mesh object, it does not create or generate a mesh
which is the work of high level mesh generation routines.

\item[]\textbf{\textit{int} MESH\_InitFromFile(\textit{Mesh\_ptr} mesh,
const char *filename):} Initialize or read a mesh from file into the
given mesh object. Returns 1 if successful, 0 otherwise.

\item[]\textbf{\textit{void} MESH\_WriteToFile(\textit{Mesh\_ptr} mesh,
const char *filename):} Save a mesh to a filename. The file is created
if it does not exists. It is recommended that the .mstk extension be
used for MSTK mesh files.  However, there is no such requirement.

\item[]\textbf{\textit{GModel\_ptr} MESH\_GModel(\textit{Mesh\_ptr} mesh):}
Return a handle to the underlying geometric model. If there is no
geometric model associated with the mesh, NULL pointer is returned.

\item[]\textbf{\textit{RepType} MESH\_RepType(\textit{Mesh\_ptr} mesh):}
Representation type currently being used by the mesh.

\item[]\textbf{\textit{int} MESH\_Num\_Vertices(\textit{Mesh\_ptr} mesh):}
Number of vertices in the mesh.

\item[]\textbf{\textit{int} MESH\_Num\_Edges(\textit{Mesh\_ptr} mesh):}
Number of edges in the mesh. For reduced representations, this routine
returns 0 since it is impractically expensive to count the number of
edges when they do not explicitly exist. Applications must find a way
to avoid using this routine for reduced representations.

\item[]\textbf{\textit{int} MESH\_Num\_Faces(\textit{Mesh\_ptr} mesh):}
Number of faces in the mesh. For reduced representations R1 or R2,
this routine counts only the faces that are explicitly represented
i.e. faces not connected to any mesh region. Therefore, a value of 0
will be returned for the number of faces of a tetrahedral mesh with
representation R1 or R2 but the correct number will be reported for a
tetrahedral mesh in other representations. Also, the correct number
will be reported for the number of faces in a surface mesh in
representation R1 or R2.  Therefore, this routine must be used
carefully.

\item[]\textbf{\textit{int} MESH\_Num\_Regions(\textit{Mesh\_ptr} mesh):}
Number of regions in the mesh.

\item[]\textbf{\textit{MVertex\_ptr} MESH\_Vertex(\textit{Mesh\_ptr} mesh,
\textit{int} i):} Return the i'th vertex in the mesh. Returns NULL if i
$<$ 0 or i $>$ number of mesh vertices.

\item[]\textbf{\textit{MEdge\_ptr} MESH\_Edge(\textit{Mesh\_ptr} mesh,
\textit{int} i):} Return the i'th edge in the mesh. Returns NULL if i $<$
0 or i $>$ number of mesh edges. Returns NULL for reduced
representations.

\item[]\textbf{\textit{MFace\_ptr} MESH\_Face(\textit{Mesh\_ptr} mesh,
\textit{int} i):} Return the i'th face in the mesh. Returns NULL if i $<$
0 or i $>$ number of mesh faces. Only faces explicitly represented in
the mesh are returned for reduced representation (See explanation for
MESH\_Num\_Faces).

\item[]\textbf{\textit{MRegion\_ptr} MESH\_Region(\textit{Mesh\_ptr} mesh,
\textit{int} i):} Return the i'th region in the mesh. Returns NULL if i
$<$ 0 or i $>$ number of mesh region.

\item[]\textbf{\textit{MVertex\_ptr} MESH\_Next\_Vertex(\textit{Mesh\_ptr}
mesh, \textit{int} *idx):} Returns the next vertex while iterating
through the vertices of the mesh. See the routine \textbf{Set\_Next\_Entry}
above for an explanation of how the iteration works.

\item[]\textbf{\textit{MEdge\_ptr} MESH\_Next\_Edge(\textit{Mesh\_ptr} mesh,
\textit{int} *idx):} Returns the next edge while iterating through the
edges of the mesh. See the routine Set\_Next\_Entry above for an
explanation of how the iteration works.  The routine always returns
NULL for reduced representations.

\item[]\textbf{\textit{MFace\_ptr} MESH\_Next\_Face(\textit{Mesh\_ptr} mesh,
\textit{int} *idx):} Returns the next face while iterating through the
faces of the mesh. See the routine \textbf{Set\_Next\_Entry} above for
an explanation of how the iteration works.  Only faces explicitly
represented in the mesh are returned for reduced representation (See
explanation for \textbf{MESH\_Num\_Faces}).

\item[]\textbf{\textit{void} MESH\_Add\_Vertex(\textit{Mesh\_ptr} mesh,
\textit{MVertex\_ptr} v):} Add a vertex to the mesh. It is assumed that
the vertex and its coordinates set are properly defined.

\item[]\textbf{\textit{void} MESH\_Add\_Edge(\textit{Mesh\_ptr} mesh,
\textit{MEdge\_ptr} e):} Add an edge to the mesh. It is assumed that
the edge is and its topology is defined.

\item[]\textbf{\textit{void} MESH\_Add\_Face(\textit{Mesh\_ptr} mesh,
\textit{MFace\_ptr} f):} Add a face to the mesh. It is assumed that the
face and its topology is properly defined.

\item[]\textbf{\textit{void} MESH\_Add\_Region(\textit{Mesh\_ptr} mesh,
\textit{MRegion\_ptr} r):} Add a region to the mesh. It is assumed that
the region and its topology is properly defined.

\item[]\textbf{\textit{void} MESH\_Rem\_Vertex(\textit{Mesh\_ptr} mesh,
\textit{MVertex\_ptr} v):} Remove vertex from mesh. Vertex is not
deleted and must be deleted afterward separately.

\item[]\textbf{\textit{void} MESH\_Rem\_Edge(\textit{Mesh\_ptr} mesh,
\textit{MEdge\_ptr} e):} Remove edge from mesh. Edge is not deleted and
must be deleted afterward separately.

\item[]\textbf{\textit{void} MESH\_Rem\_Face(\textit{Mesh\_ptr} mesh,
\textit{MFace\_ptr} f):} Remove face from mesh. Face is not deleted and
must be deleted afterward separately.

\item[]\textbf{\textit{void} MESH\_Rem\_Region(\textit{Mesh\_ptr} mesh,
\textit{MRegion\_ptr} r):} Remove region from mesh. Region is not deleted and must be deleted afterward separately.

\item[]\textbf{\textit{void} MESH\_Set\_GModel(\textit{Mesh\_ptr} mesh,
GModel\_ptr geom):} Assign a geometric model handle to the mesh.

\item[]\textbf{\textit{int} MESH\_Change\_RepType(\textit{Mesh\_ptr} mesh,
\textit{int} nurep):} Change the representation type of the mesh. This
routine can be used to modify the representation type dynamically to
suit different algorithms.  However, the cost of making the change and
reordering all adjacencies and creating or deleting entities has to be
considered while invoking this routine. Also, once a conversion is
made from a full representation to a reduced representation, not all
information may be retrievable when switching back to a full
representation.  (particularly classification information, i.e.,
relationship of mesh entities to the geometric model).

\end{description}

\newpage
\subsection{Mesh Vertex Object}

\begin{description}
\item[]\textbf{\textit{MVertex\_ptr} MV\_New(\textit{Mesh\_ptr} mesh):}
Create a new vertex object. No geometric or topological information is
embedded in the vertex when it is created. The vertex only knows which
mesh it belongs to. The ID of the vertex is set by this function.

\item[]\textbf{\textit{void} MV\_Delete(\textit{MVertex\_ptr} mvertex):}
Delete the vertex. Deletes all topological and geometric
information embedded in the vertex.

\item[]\textbf{\textit{void} MV\_Set\_Coords(\textit{MVertex\_ptr} mvertex,
double *xyz):} Set the coordinates of the vertex.

\item[]\textbf{\textit{void} MV\_Set\_GEntity(\textit{MVertex\_ptr} mvertex,
GEntity\_ptr gent):} Set the geometric model entity on which vertex is
classified.

\item[]\textbf{\textit{void} MV\_Set\_GEntDim(\textit{MVertex\_ptr} mvertex,
\textit{int} gdim):} Set topological dimension of model entity on which
vertex is classified.

\item[]\textbf{\textit{void} MV\_Set\_GEntID(\textit{MVertex\_ptr} mvertex,
\textit{int} gid):} Set ID of model entity on which vertex is
classified.

\item[]\textbf{\textit{void} MV\_Add\_AdjVertex(\textit{MVertex\_ptr}
mvertex, \textit{MVertex\_ptr} adjvertex):} Add neighboring vertex,
adjvertex, to ajdacent vertex list of vertex, mvertex.

\item[]\textbf{\textit{void} MV\_Rem\_AdjVertex(\textit{MVertex\_ptr}
mvertex, \textit{MVertex\_ptr} adjvertex):} Delete neighboring vertex
of given vertex.

\item[]\textbf{\textit{void} MV\_Set\_ID(\textit{MVertex\_ptr} mvertex,
\textit{int} id):} Explicitly set ID of a vertex and overwrite the ID
set by the MV\_New operator. Does not check for duplication of edge
IDs.

\item[]\textbf{\textit{Mesh\_ptr} MV\_Mesh(\textit{MVertex\_ptr} mv):}
Returns the mesh that this vertex belongs to.

\item[]\textbf{\textit{int} MV\_ID(\textit{MVertex\_ptr} mvertex):} Returns
the ID of the vertex.

\item[]\textbf{\textit{int} MV\_GEntDim(\textit{MVertex\_ptr} mvertex):}
Returns the dimension of the geometric model entity that the vertex is
classified on. Returns -1 if not known.

\item[]\textbf{\textit{int} MV\_GEntID(\textit{MVertex\_ptr} mvertex):}
Returns the ID of the geometric model entity that the vertex is
classified on. Returns 0 if this information is not known.

\item[]\textbf{\textit{GEntity\_ptr} MV\_GEntity(\textit{MVertex\_ptr}
  mvertex):} Returns a pointer or handle to the geometric model entity
  that the vertex is classified on. Returns NULL if this information
  is not known.

\item[]\textbf{\textit{void} MV\_Coords(\textit{MVertex\_ptr} mvertex,
double *xyz):} Returns the coordinates of the vertex.

\item[]\textbf{\textit{int} MV\_Num\_AdjVertices(\textit{MVertex\_ptr}
mvertex):} Returns the number of edge connected neighboring vertices of
vertex. \textit{Not efficient for all representations}.

\item[]\textbf{\textit{int} MV\_Num\_Edges(\textit{MVertex\_ptr} mvertex):}
Returns the number of edges connected to the vertex.

\item[]\textbf{\textit{int} MV\_Num\_Faces(\textit{MVertex\_ptr} mvertex):}
Returns the number of faces connected to the vertex.

\item[]\textbf{\textit{int} MV\_Num\_Regions(\textit{MVertex\_ptr} mvertex):}
Returns the number of regions connected to the vertex

\item[]\textbf{\textit{Set\_ptr} MV\_AdjVertices(\textit{MVertex\_ptr}
mvertex):} Set of adjacent or edge connected neighboring vertices of vertex.

\item[]\textbf{\textit{Set\_ptr} MV\_Edges(\textit{MVertex\_ptr} mvertex):}
Set of edges connected to the vertex.

\item[]\textbf{\textit{Set\_ptr} MV\_Faces(\textit{MVertex\_ptr} mvertex):}
Set of faces connected to the vertex.

\item[]\textbf{\textit{Set\_ptr} MV\_Regions(\textit{MVertex\_ptr} mvertex):}
Set of regions connected to the vertex.
\end{description}

\newpage
\subsection{Mesh Edge Object}

\begin{description}
\item[]\textbf{\textit{MEdge\_ptr} ME\_New(\textit{Mesh\_ptr} mesh):} Create
a new edge object. No topological information is embedded in the edge
when it is created. The edge only knows which mesh it belongs
to. The ID of the edge is set by this function.

\item[]\textbf{\textit{void} ME\_Delete(\textit{MEdge\_ptr} medge):} Delete
the edge. Deletes all topological information embedded in the edge.

\item[]\textbf{\textit{void} ME\_Set\_GEntity(\textit{MEdge\_ptr} medge,
GEntity\_ptr gent):} Set the geometric model entity on which the edge
is classified.

\item[]\textbf{\textit{void} ME\_Set\_GEntDim(\textit{MEdge\_ptr} medge,
\textit{int} gdim):} Set the topological dimension of model entity on
which edge is classified.

\item[]\textbf{\textit{void} ME\_Set\_GEntID(\textit{MEdge\_ptr} medge,
\textit{int} gid):} Set ID of model entity on which edge is classified.

\item[]\textbf{\textit{void} ME\_Set\_ID(\textit{MEdge\_ptr} medge,
\textit{int} id):} Explicitly set ID of an edge and overwrite the ID set
by the ME\_New function. Does not check for duplication of edge IDs.

\item[]\textbf{\textit{void} ME\_Set\_Vertex(\textit{MEdge\_ptr} medge,
\textit{int} i, \textit{MVertex\_ptr} vertex):} Set the i'th vertex of
the edge. i can be 0 or 1.

\item[]\textbf{\textit{void} ME\_Replace\_Vertex(\textit{MEdge\_ptr} medge,
\textit{MVertex\_ptr} vert, \textit{MVertex\_ptr} nuvert):} Replace
i'th vertex by new vertex.

\item[]\textbf{\textit{Mesh\_ptr} ME\_Mesh(\textit{MEdge\_ptr} medge):}
Returns the mesh that this edge belongs to.

\item[]\textbf{\textit{int} ME\_ID(\textit{MEdge\_ptr} medge):} Returns the
ID of the vertex. Returns -1 if not known.

\item[]\textbf{\textit{int} ME\_GEntDim(\textit{MEdge\_ptr} medge):} Returns
the dimension of the geometric model entity that the vertex is
classified on. Returns -1 if not known.

\item[]\textbf{\textit{int} ME\_GEntID(\textit{MEdge\_ptr} medge):} Returns
the ID of the geometric model entity that the vertex is classified on.
Returns 0 if this information is not known.

\item[]\textbf{\textit{GEntity\_ptr} ME\_GEntity(\textit{MEdge\_ptr} medge):} Returns
a pointer or handle to the geometric model entity that the vertex is
classified on. Returns NULL if this information is not known.

\item[]\textbf{\textit{int} ME\_Num\_Faces(\textit{MEdge\_ptr} medge):}
Returns the number of faces connected to the edge.

\item[]\textbf{\textit{int} ME\_Num\_Regions(\textit{MEdge\_ptr} medge):}
Returns the number of regions connected to the edge.

\item[]\textbf{\textit{MVertex\_ptr} ME\_Vertex(\textit{MEdge\_ptr} medge,
\textit{int} i):} Returns the i'th vertex of the edge. i=0 returns the
first vertex and i=1 returns the second vertex.

\item[]\textbf{\textit{MVertex\_ptr} ME\_OppVertex(\textit{MEdge\_ptr}
medge, \textit{MVertex\_ptr} ov):} Return the vertex opposite to given
vertex in edge.

\item[]\textbf{\textit{int} ME\_UsesEntity(\textit{MEdge\_ptr} medge,
\textit{MEntity\_ptr} mentity, \textit{int} etype):} Check if edge uses
given lower dimension entity, \textit{mentity}. The dimension of the
entity is specified by the \textit{etype} variable. For an edge, the
only lower dimensional entity is a vertex. If the edge uses the
vertex, the function returns 1; otherwise it returns 0. If any
other type of entity is specified, the function returns 0.

\item[]\textbf{\textit{Set\_ptr} ME\_Faces(\textit{MEdge\_ptr} medge):}
Returns the set of faces using this edge.

\item[]\textbf{\textit{Set\_ptr} ME\_Regions(\textit{MEdge\_ptr} medge):}
Returns the set of regions using this edge.

\item[]\textbf{\textit{MEdge\_ptr} MVs\_CommonEdge(\textit{MVertex\_ptr} v1,
\textit{MVertex\_ptr} v2):} Return the edge connecting vertices v1 and
v2, if it exists. If such an edge does not exist, the function returns
0.

\item[]\textbf{\textit{double} ME\_Len(\textit{MEdge\_ptr} e):} Return
  the length of the straight line connecting the two vertices of the
  edge.

\item[]\textbf{\textit{double} ME\_LenSqr(\textit{MEdge\_ptr} e):}
  Return the square of the length of the straight line connecting the
  two vertices of the edge.

\item[]\textbf{\textit{void} ME\_Vec(\textit{MEdge\_ptr} e, double *evec):}
Return the vector going from the first vertex of the edge to the
second vertex of the edge.
\end{description}

\newpage
\subsection{Mesh Face Object}

\begin{description}
\item[]\textbf{\textit{MFace\_ptr} MF\_New(\textit{Mesh\_ptr} mesh):} Create
a new face object. No topological information is embedded in the face
when it is created. The face only knows which mesh it belongs
to. The ID of the face is set by this function.

\item[]\textbf{\textit{void} MF\_Delete(\textit{MFace\_ptr} mface):} Delete
the face. Delete all topological information embedded in the face.

\item[]\textbf{\textit{void} MF\_Set\_GEntity(\textit{MFace\_ptr} mface, GEntity\_ptr gent):} Set the geometric model entity on which the edge is classified.

\item[]\textbf{\textit{void} MF\_Set\_GEntDim(\textit{MFace\_ptr} mface,
\textit{int} gdim):} Set the dimension of the geometric model entity on
which the face is classified.

\item[]\textbf{\textit{void} MF\_Set\_GEntID(\textit{MFace\_ptr} mface,
\textit{int} gid):} Set the ID of the geometric model entity on which
the face is classified.

\item[]\textbf{\textit{void} MF\_Set\_ID(\textit{MFace\_ptr} mface,
\textit{int} id):} Explicitly set ID of an edge and overwrite the ID
set by the MF\_New operator. Does not check for duplication of face
IDs.

\item[]\textbf{\textit{void} MF\_Set\_Edges(\textit{MFace\_ptr} mface,
\textit{int} n, \textit{MEdge\_ptr} *edges, \textit{int} *dirs):} Set
the edges of the face along with their directions. The ordered set of
edge pointers and their directions are passed in through arrays along
with the number of edges. The edges are assumed to be ordered
clockwise around the face. If an edge direction is along the clockwise
direction of the face then the entry in the 'dirs' array must be 1;
otherwise it must be 0. This function is relevant only for full
representations in MSTK.

\item[]\textbf{\textit{void} MF\_Set\_Vertices(\textit{MFace\_ptr} mface,
\textit{int} n, \textit{MVertex\_ptr} *verts):} Set the vertices of the
face. The ordered set of vertices is passed in through an array along
with the number of vertices. The vertices are assumed to be ordered
clockwise around the face. This function is relevant only for reduced
representations in MSTK.

\item[]\textbf{\textit{void} MF\_Replace\_Edge(\textit{MFace\_ptr} mface,
\textit{MEdge\_ptr} edge, \textit{MEdge\_ptr} nuedge, \textit{int}
dir):} Replace an edge in the face with another edge. The direction
in which the new edge is used in the face must also be supplied. This
function is relevant only for full representations in MSTK.

\item[]\textbf{\textit{void} MF\_Replace\_Vertex(\textit{MFace\_ptr} mface,
\textit{MVertex\_ptr} mvertex, \textit{MVertex\_ptr} nuvertex):}
Replace a vertex in the face with another vertex. This function is
relevant only for reduced representations in MSTK.

\item[]\textbf{\textit{void} MF\_Replace\_Edge\_i(\textit{MFace\_ptr} mface,
\textit{int} i, \textit{MEdge\_ptr} nuedge, \textit{int} dir):} Replace
the i'th edge in the face with another edge. The direction in which the
new edge is used in the face must also be supplied. This function is
relevant only for full representations in MSTK.

\item[]\textbf{\textit{void} MF\_Replace\_Vertex\_i(\textit{MFace\_ptr}
mface, \textit{int} i, \textit{MVertex\_ptr} nuvertex):} Replace the
i'th vertex in the face with a new vertex. This function is relevant
only for reduced representations in MSTK.

\item[]\textbf{\textit{Mesh\_ptr} MF\_Mesh(\textit{MFace\_ptr} mf):} Returns the mesh that this mesh belongs to.

\item[]\textbf{\textit{int} MF\_ID(\textit{MFace\_ptr} mface):} Returns the ID of the face. Returns 0 if not known.

\item[]\textbf{\textit{int} MF\_GEntDim(\textit{MFace\_ptr} mface):} Returns the dimensions of the geometric model entity that the vertex is classified on. Returns -1 if not known.

\item[]\textbf{\textit{int} MF\_GEntID(\textit{MFace\_ptr} mface):} Returns the ID of the geometric model entity that the vertex is classified on. Returns 0 if this information is not known.

\item[]\textbf{\textit{GEntity\_ptr} MF\_GEntity(\textit{MFace\_ptr} mface):} Returns a pointer or handle to the geometric model entity that the vertex is classified on. Returns NULL if this information is not known.

\item[]\textbf{\textit{int} MF\_Num\_Vertices(\textit{MFace\_ptr} mface):} Returns the number of vertices of the face.

\item[]\textbf{\textit{int} MF\_Num\_Edges(\textit{MFace\_ptr} mface):} Returns the number of edges of the face.

\item[]\textbf{\textit{int} MF\_Num\_AdjFaces(\textit{MFace\_ptr}
    mface):} Returns the number of adjacent faces of a face. This
  operator is relevant only in planar or surface meshes, i.e., for
  boundary faces not connected to any regions.

\item[]\textbf{\textit{Set\_ptr} MF\_Vertices(\textit{MFace\_ptr}
    mface, \textit{int} dir, \textit{MVertex\_ptr} mvert):} Return the
  ordered set of the vertices of the face. The vertices are ordered in
  ccw direction while looking down the face 'normal', if 'dir' is 1
  and in the cw direction, if 'dir' is 0. If 'mvert' is specified, the
  vertex set is reordered so that it is the first vertex (This
  argument will be added soon to the function. For now, omit this
  argument). The behavior of this function can be illustrated using
  Figure~\ref{fig:face_def}. For the face shown in the figure, a
  vertex set with ccw ordering or 'dir' = 1 is ${V_0,V_1,V_2,V_3}$ and
  a vertex set with cw ordering or 'dir' = 0 is ${V_0,V_3,V_2,V_1}$.
  A vertex set with ccw ordering starting with vertex $V_2$ is
  ${V_2,V_3,V_0,V_1}$.

\begin{figure}[!h]
  \begin{center}
  \includegraphics[scale=1.0]{figures/face_def} 
  \caption{Face definition}
  \label{fig:face_def}
  \end{center}
\end{figure}

\item[]\textbf{\textit{Set\_ptr} MF\_Edges(\textit{MFace\_ptr} mface,
    \textit{int} dir, \textit{MVertex\_ptr} mvert):} Return the
  ordered set of edges of the face. The edges are ordered in the ccw
  while looking down the face 'normal' if dir is 1 and in the cw if
  dir is 0. If 'mvert' is specified, the edge set is reordered so that
  the first edge in the set contains this vertex. More precisely, if
  'dir' is 1, and the first edge is 'e' used in the face in direction
  'd', then \textbf{ME\_Vertex(e,!d)} = mvert. With reference to
  Figure~\ref{fig:face_def}, the edges of the face in the ccw direction or
  'dir' = 1 are ${E_0,E_1,E_2,E_3}$ and in the opposite dir are
  ${E_3,E_2,E_1,E_0}$. If 'mvert' or the starting vertex is specified
  as $V_2$, the edge set in ccw direction or 'dir' = 1 is
  ${E_2,E_3,E_0,E_1}$ and in the opposite direction is
  ${E_1,E_0,E_3,E_2}$.

\item[]\textbf{\textit{Set\_ptr} MF\_AdjFaces(\textit{MFace\_ptr} mface):}
Set of adacent faces of a face. This operator is relevant only in
planar or surface surface meshes, i.e., for boundary faces not
connected to any regions.

\item[]\textbf{\textit{int} MF\_EdgeDir(\textit{MFace\_ptr} mface,
\textit{MEdge\_ptr} medge):} Returns the direction in which the face
uses the given edge. If the faces use the edge in the positive direction,
the function returns 1; otherwise it returns 0.

\item[]\textbf{\textit{int} MF\_EdgeDir\_i(\textit{MFace\_ptr} mface,
\textit{int} i):} Returns the direction in which the face uses its i'th
edge. If the face uses the edge in the positive direction the function
returns 1; otherwise it returns 0;

\item[]\textbf{\textit{int} MF\_UsesEntity(\textit{MFace\_ptr} mface,
\textit{MEntity\_ptr} mentity, \textit{int} type):} Check if the face
uses the given lower dimension entity, 'mentity'. The type of the
entity is specified by the 'etype' variable. For a face, a lower
dimensional entity can be a vertex or an edge. If the face uses the
vertex or the edge, the function returns 1; otherwise it returns 0. If
any other type of entity is specified, the function returns 0.

\item[]\textbf{\textit{Set\_ptr} MF\_Regions(\textit{MFace\_ptr} mface):}
Return the set of regions connected to the face. If the face is not
used by any regions, the function returns NULL to indicate that the
set is empty. If not the set may contain one or two regions.

\item[]\textbf{\textit{MRegion\_ptr} MF\_Region(\textit{MFace\_ptr} mface,
\textit{int} side):} Returns the region on the specified side of the
face. The positive side of the face (side = 1) is the side towards
which the face normal points. The negative side of the face (side = 0)
is the opposite side.

\item[]\textbf{\textit{void} MF\_Coords(\textit{MFace\_ptr} mface,
\textit{int} *n, double (*xyz)[3], \textit{int} dir):} Returns the
coordinates of the face vertices in an array along with the number of
vertices. If 'dir' is 1, the coordinates are returned with a  ccw
ordering while looking down the face normal; if 'dir' is 0, they are
returned with a cw ordering (See Figure~\ref{fig:face_def}).
\end{description}

\newpage
\subsection{Mesh Region Object}

\begin{description}
\item[]\textbf{\textit{MRegion\_ptr} MR\_New(\textit{Mesh\_ptr} mesh):}
Create a new region object. No topological information is embedded in
the region when it is created. The region only knows which mesh it
belongs to. The ID of the region is set by this function.

\item[]\textbf{\textit{void} MR\_Delete(\textit{MRegion\_ptr} mregion):}
Delete the region. Deletes all topological information embedded in the
edge.

\item[]\textbf{\textit{void} MR\_Set\_GEntity(\textit{MRegion\_ptr} mregion,
GEntity\_ptr gent):} Set the geometric model entity on which the edge
is classified.

\item[]\textbf{\textit{void} MR\_Set\_GEntDim(\textit{MRegion\_ptr} mregion,
\textit{int} gdim):} Set the dimension of the geometric model entity on
which the edge is classified.

\item[]\textbf{\textit{void} MR\_Set\_GEntID(\textit{MRegion\_ptr} mregion,
\textit{int} gid):} Set the ID of the geometric model entity on which
the edge is classified.

\item[]\textbf{\textit{void} MR\_Set\_ID(\textit{MRegion\_ptr} mregion,
\textit{int} id):} Explicitly set ID of an edge and overwrite the ID
set by the \textbf{MR\_New} function. Does not check for duplication
of region IDs.

\item[]\textbf{\textit{void} MR\_Set\_Faces(\textit{MRegion\_ptr} mregion,
\textit{int} nf, \textit{MFace\_ptr} *mfaces, \textit{int} *dirs):} Set
the faces of the region along with their directions. The
\textit{unordered} set of faces and their directions are passed in
through arrays along with the number of faces. If the normal of the
face points out of the region, the associated direction to be passed
in is 1; otherwise it is 0. This function is only relevant for full
representations in MSTK.

\item[]\textbf{\textit{void} MR\_Set\_Vertices(\textit{MRegion\_ptr}
mregion, \textit{int} nv, \textit{MVertex\_ptr} *mvertices):} Set the
vertices of the region. This function is relevant for reduced
representations only. For standard elements, the vertices must be
ordered as indicated in Appendix~\ref{app:reg_conventions}. This
function does not apply for defining general regions since there
cannot be any implicit ordering.

\item[]\textbf{\textit{void} MR\_Replace\_Face(\textit{MRegion\_ptr}
mregion, \textit{MFace\_ptr} mface, \textit{MFace\_ptr} nuface,
\textit{int} dir):} Replace a face of the region with another edge. The
direction in which the new face is used in the region must also be
supplied. This function is only relevant for full representations in
MSTK.
